Discovery of Rotational Modulations in the Planetary-mass Companion 2M1207\lowercase{b}: An intermediate Rotation Period and Heterogeneous Clouds in a Low Gravity Atmosphere

### 2015-11-01 summary/referee_report_reply.md

# Referee report reply

## Major comment:

> At the beginning of Sec 4 the authors claim that the photometry 'showed apparently sinusoidal modulations'. They proceed to fit sinusoids to the light curve to determine a period and find periods between 9 and 10 hours. I would argue that this is a dubious approach. The first sensible step is to actually prove that a period is present in the data (in addition to the subjective eye test), before modeling the light curve with a sinusoid.

>As argument in favor of their period detection the authors say that the fitted periods are 'very different from all timescales present in our observations' - that is true, except for ONE timescale, the total duration of the observations, which is around 9 hours. Playing devil's advocate, fitting a sinusoid to a light curve with just a linear trend tends to result in a period around the total length of the dataset. Looking at the light curves, it is not so evident to me that a periodic curve is necessarily the best model - a linear fit would make much sense too, particularly when the first orbit is discarded (as the authors argue, the 1st orbit is often unstable). In that case, 10 hours could be the lower limit of the rotation period.

>What I would like to see is a statistical test that confirms that a periodicity is in fact a better fit to the data than a linear trend. Alternatively, the authors could use some of the well established period search routines to bolster the evidence for their detection. Since the rotational modulation is the key result of this paper, it seems prudent to me to not only rely on the eye test, in particular since there is a wealth of analysis routines available for this kind of problem.

>Parts of introduction, discussion, and abstract may need to be revised after verifying the presumed period.
*****
>Sec 4.2: I do not fully understand why the authors add noise to their light curve and THEN fit sinusoids. Why is the noise increased artificially for this procedure? This seems to be a very unusual period finding algorithm.

The minor comment on the period finding algorithm is directly related to the major comment. Therefore we combine our reply to these comments together.

We agree with the referee that our original fitting method is inappropriate in the way that it artificially adds noise to the light curve. Therefore we did a rigorous MCMC fitting to measure the period and amplitude based on a sinusoidal model. We cannot use normal period search routines (e.g. Fourier analysis, auto-correlation) because the period of the modulation is larger than our observation baseline. We assumed uniform distribution for the period, amplitude and phase of the sinusoid as priors and included period as long as 40 hours as possible solution. We did the fitting for both including and excluding data from the first orbit. Even when the data from the first orbit were excluded, the posterior distributions for the amplitude and period were constraint, though not as well as those for the fitting include the whole dataset. Our measurements for period and amplitude agree within 1-sigma uncertainties with or without using data from the first orbit. Using the posterior distribution of MCMC fitting, we can show the statistical significance of our measurement.

We also added discussion from the perspective of physical viability of the light curve models and explained a sinusoid is the most suitable choice. A complete flat line is a viable physical that would mean no variability is detected. A flat line with non-zero slope is not a physically viable model: we know that 2M1207b is not fainting away or brightening indefinitely. Therefore, a model that is periodic or irregularly variable (but with a constant mean) is a physical model. The simplest of these physically viable non-constant brightness models is a sine wave.

Our fits demonstrate that the constant level (flat line) can be excluded. Of the non-constant physically viable models the sine wave gives a reasonable chi2, but more complex models are not warranted by our data.

We note that, as the referee also pointed out, there is no timescale linked to the observations that would introduce a 9-hour period slope. The fact that the primary’s light curve is constant to a very high accuracy also demonstrates that there is no instrumental sensitivity drift or any similar effect.

The changes of the manuscript include Sec. 4.2 re-writed, substitution of Fig. 4 with MCMC posterior distribution, and expanding the first paragraph of Sec. 6 for above discussion of physical viability.

## Minor comments:

>- Sec 1, end of 2nd paragraph: It might help to say 'comparing to isolated brown dwarfs', to make the sentence clearer.

We added word *isolated* for clarity.

>- Sec 1, end of 3rd paragraph: For the formation scenarios for the 2M1207 system, Lodato et al. (2005) is an important reference.

We added the reference Lodato et. al. (2005)

>- Sec 1, penultimate paragraph: I would like to see some references for the 'discovery of additional planetary-mass companions' (although they are given earlier in this section).

References for AB pic b and HR8799bcde were added as examples of additional PMC that has red color and under-luminosity problems

>- Sec 2: This is the part that might be a little too concise. One basic information I am missing is the exposure time. It would also help to point out that 1.5 minutes is the MAXIMUM cadence (if I understand this correctly). Also, 'we took 8 SPARS10 exposure sequence' might need a bit more information to help the reader.

We added information for Sec. 2. Exposure time information included, SPARS10, NSAMP=10 are explained. Major modification is the following sentence:

**Original**:In each orbit we took 8 SPARS10 exposure sequence with NSAMP=10, alternating between F160W and F125W filters, with 2--3 identical exposures in each exposure sequence.

**Revised** In each orbit, we took 8 SPARS 10 exposure sequences alternating between F125W and F160W filters. Each sequence contained 2--3 identical exposures that had 10 non-destructive read-outs and total exposure time of 88.4 s.

We added word *maximum* before cadence.

>- Sect 3: I like the innovative approach for modeling the PSF, but I would also like to see this demonstrated in a figure, i.e. an extension of Fig. 1 showing the original image, the correction image, and the two PSFs.

We included the examples of Tiny Tim PSFs as well as the correction map and modified the figure caption correspondingly.

>- Fig 2+3: Information about the epoch of the observations in JD needs to be added, either here or in the text.

JD is added at the beginning of Section 2.

>- Sec 6, 2nd paragraph: For disk masses, Scholz et al. (2006) is an important
additional reference.

Reference Scholz et. al. (2006) added for disk masses.

>- The discussion in Sec 6 touches on many interesting points, but unfortunately remains relatively vague. At the end of the 2nd paragraph, the authors write 'may place powerful constraints on the evolution of the internal structures of these contracting objects'. Maybe it is worth saying a little more here? What exactly is expected to change in the internal structures of these fully convective objects?

>- For the discussion of the rotational evolution of brown dwarfs, two relevant references that are missing are the PPVI review by Bouvier et al. (see Sect. 4.3.) and, as an update, the recent paper by Scholz, Kostov, et al.. This second paper is important because it contains a large sample of rotation periods for brown dwarfs at about the same age as 2M1207 (although most of them are more massive than 2M1207). At this age some effect of disk braking is still observed, while for the continuing evolution wind braking does play a role in the substellar regime. If 2M1207b really has a rotation period of 10h, it would be a planetary-mass counterpart of the fastest rotating young brown dwarfs. (It would also rotate at a significant fraction of breakup velocity, which might also be of interest.)

We agree with the referee at this point. We expanded the discussion from the point of view of the evolution of angular momentum mainly based on the two references recommended by the referee. We calculated the break-up rotation period is ~2 hr assuming a mass of 5 M_jup and a radius of 1.4 R_jup. The measured rotation period ~10 hr indicates rotation rate is much smaller than the break-up rotation limit.

Scholz et al. 2015 measures rotation period of 20 young brown dwarfs, and compares the BD rotation period measurements from 3 star formation region with rotation evolution tracks based on conservation of angular momentum. The age of brown dwarf in this sample is similar to 2M1207b (~6-9Myr) and the rotation periods of the brown dwarfs in this sample is closer to 2M1207b comparing to Metchev et al. (2015)

We added one paragraph to discuss the comparison of rotation period with BDs, and the evolution of angular momentum of objects with fully convective interior:

**Rotation period of ∼ 10 hour is significantly larger than field brown dwarfs from the sample of Metchev et al. (2015), but comparable to the me- dian period of the sample of Scholz et al. (2015), of which the age is similar to 2M1207b. The rotation period of 2M1207b is agreed within the range predicted by evolutionary track established from measurements of brown dwarfs assuming conservation of angular momentum (Fig. 4 in Scholz et al. (2015)), and much longer than the break-up limit, i.e. the rotation period where the equatorial centrifugal force exceeded gravitational force. Observed rotation rate for brown dwarfs of different ages show little evidence of angular momentum loss at the age of several Myr as oppose to low mass stars (Bouvier et al. 2014; Scholz et al. 2015), and agree with the model of solid body rotation for the fully convective interior. Rotation periods of planetary mass objects with well established age measurement, can place stronger constrain on the internal structure, since angular momentum loss through disks is even less likely comparing with brown dwarfs.**

We added the measurements from Scholz 2015 to left panel of Figure 5.

>- Sec 7: 'significantly longer than most field BDs with known rotation period'
>- I would advise to be careful with this statement. The available sample of field BD rotation periods is certainly very biased. The vsini data wouldindicate that rotation periods of 10-20 hours may not be that unusual (Reiners& Basri 2008).

Considering the period measurement from sample of Scholz et. al. 2015, We agree with referee at this point.

We modified following sentences in the conclusion.

The period is 10.5 hours, similar to that derived from v sin i measurements for the directly imaged exoplanet β Pic b, significantly longer than most field brown dwarfs with known rotation periods, **but comparable to the brown dwarfs in a sample with similar age to 2M1207b.**

*****

**We include two additional references.**

1. Biller et. al. (2015) (arXiv1510.07625) recently published time resolved observation on PSO J318.5-22, an isolated planetary mass object, and detected variability. We added the reference as an example of time resolved observation and rotational mapping.
2. Karalidi et. al. (2015) published a rotational mapping code that retrieve 2D atmospheric structure from light curves. We added one paragraph at the end of discussion section about the possible application of this tool in the future.

### 2015-11-01 summary/letter/letter_vAG.tex
revision finished

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\newcommand{\tinytim}{\textit{Tiny Tim}}
\newcommand{\bpic}{$\beta$ Pic}
\newcommand{\vsini}{$v\sin i$}
\newcommand{\mjup}{M$_{\mbox{Jup}}$}
\begin{document}

@@

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, periodic light curves directly probe the
  rotational periods of ultra-cool objects.  We present here, for the
  first time, time-resolved high-precision photometric measurements of
  a planetary-mass companion, 2M1207b.  We observed the binary
  system with HST/WFC3 in two bands and with two spacecraft roll
  angles. Using point spread function-based photometry, we reach a
  nearly photon-noise limited accuracy for both the primary and the
  secondary. While the primary is consistent with a flat light curve,
  the secondary shows modulations that are clearly detected in the
  combined light curve as well as in different subsets of the data.
  The amplitudes are 1.45\% in the F125W and 0.92\% in the F160W
  filters, respectively. We find a consistent period of
  $10.2^{+0.9}_{-0.8}$ hours and
  similar phases in both bands. The J- and H-band amplitude ratio of
  2M1207b is very similar to a field brown dwarf that has identical
  spectral type but different J-H color.  Importantly, our study also
  measures, for the first time, the rotation period for a directly 
  imaged planetary-mass companion to another star.
\end{abstract}

@@

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}
  \caption{Point spread function subtraction allows isolating the
    secondary and accurately measuring its brightness in our WFC3
    F160W images. {\em Upper}: original image, the position of 2M1207
    b is indicated by a white circle.  {\em Lower:} residual image --
    after the subtraction of the hybrid PSF, 2M1207b is detected at a
    high significance level.}
  \label{fig:1}
\end{figure*}

@@

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping \citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2015,Metchev2015,Heinze2015}. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
. \citet{Kostov2013}
demonstrated that these techniques can be applied to directly imaged
exoplanets, too, allowing comparative studies of objects with
different surface gravities. However, high contrast magnifies the
challenges for directly imaged exoplanets and planetary-mass
companions to acquire high-precision light curves comparing to brown
dwarfs.

2M1207b \citet{Chauvin2004} is the first directly imaged extra-solar
planetary-mass companion. \citet{Chauvin2005} and \citet{Song2006}
confirmed that 2M1207b and its host 2M1207A form a bound, co-moving
system. 2M1207A and b have an angular separation of $0.78''$, which
corresponds to a projected separation of 41.2 AU at a distance of 52.4
pc \citep[e.g.][]{Ducourant2008}. Combining 2M1207b's age and near
infrared luminosity with brown dwarf cooling models
\citep[e.g.][]{Baraffe2003}, the object's mass is estimated to be
2.3-4.8 M$_{\mathrm{Jup}}$ \citep{Barman2011b}. Even though a
circumsubstellar disk was discovered around 2M1207A
\citep[][]{Sterzik2004}, the high companion-to-host mass ratio and
large separation argue for binary-like gravitational fragmentation
formation\citep{Mohanty2007}.

@@

The discovery of additional planetary-mass companions with similarly
red colors and apparent under-luminosity have highlighted 2M1207b as a
template of low gravity ultra-cool atmospheres but as of now
understanding the composition and structure of clouds and their
gravity-dependence remained elusive.

In this {\em Letter} we present the first, high-contrast, high-cadence,
high-precision, time-resolved {\em Hubble Space Telescope} (HST)
photometric time series of 2M1207b, a directly imaged planetary-mass
object. We successfully detect rotational modulation and measure the
amplitudes in two bands and determine the rotational period. These
observations probe the spatial heterogeneity and vertical structure of
clouds in planetary mass objects for the first time.

@@

We obtained direct images of the 2M1207A+b system on UT 2014 April 11
from 08:07:47 to 16:53:18 using HST and its Wide Field Camera 3
\citep[WFC3, pixel scale=$0.130$mas/pixel, ][]{Mackenty2008} in the frame of
the HST Program GO-13418 (PI: D. Apai). We acquired the observations
in filters F125W ($\lambda_{\mathrm{pivot}}$ = 1245.9 nm, full width
at half maximum (FWHM) = 301.5 nm) and F160W
($\lambda_{\mathrm{pivot}}$ 1540.52, FWHM = 287.9 nm), roughly
corresponding to the J and H bands. We used the $256\times256$ pixels
sub-array mode to avoid memory dumps during the observations.  In
order to provide a near-continuous coverage for detecting modulations
we observed the 2M1207 system in 6 consecutive HST orbits, obtaining
data with cadence of $\sim1.5$ minutes over a baseline of 8 hours and
40 minutes. The observations were interrupted by 58-minute long Earth
occultations every 94 minutes.

The observations applied space craft rolls each two orbits to allow
roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbit 1, 3, and 5, and those for 2, 4, and 6
differed by $25^{\circ}$. At the separation of 2M1207b, this angle
difference corresponds to a displacement of $0.34''$, or 2.75 and 2.30
resolution elements in F125W and F160W, respectively. In each orbit we
took 8 SPARS10 exposure sequence with NSAMP=10, alternating between
F160W and F125W filters, with 2--3 identical exposures in each
exposure sequence. To improve sampling and reduce the risk that the
core of point spread function (PSF) is affected by bad pixels, we
applied a 4-point dither pattern with differential ``X/Y'' offsets of
1.375" in the detector frame, providing optimal non-integral (half
pixel) step of 10.5 and 8.5 pixels in F125W and F160W,
respectively. In total, we obtained 70 and 64 images with 10 non-destructive
read-outs in F125W and F160W, respectively.

@@

@@

@@

We used a MC method to analyze the light curve, and provided amplitudes
and periods as well as their uncertainties for both filters. We
generated series of Gaussian distributed random noise with standard deviation
same as the photon noise, added the noise to the original light curves, and
fitted sinusoids to the newly combined light curves. We repeated these
procedures  for 100,000 times and obtain the distributions of period
and amplitude (Figure \ref{fig:4}).

\section{Result}
\label{Results}

\begin{figure*}
  \centering
  \includegraphics[width=0.45\textwidth]{periodDistr}
  \includegraphics[width=0.45\textwidth]{amplitudeDistr}
  \caption{Distributions for period (left) and amplitude (right) for the light
    curves of F125W and F160W. The bin sizes for the histograms of
    period  and amplitude are
  0.25 hours and 0.5\%, respectively. Histograms are
  normalized to the maxima. In the right panel, Gaussian profiles are fitted to the
  histograms and plotted in solid lines.}
  \label{fig:4}
\end{figure*}

We present the first high-contrast, high-cadence, and high-precision
photometry of a directly imaged planet or planetary-mass
companion around another star. Our observations reveal a modulation in the light curve of
the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary-mass objects.  The best
fit periods for F125W and F160W are $10.5_{-1.2}^{+1.3}$ and $9.1_{-1.0}^{+1.1}$ hours,
respectively. We also fitted the two band light curves together forcing the
periods of two sinusoids to be the same, and obtained a period
of $10.2^{+0.9}_{-0.8}$ hours.

@@

\section{Discussion}

\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change
    with brown dwarfs, \bpic{} b, and solar system planets. {\em
      Left}: period vs. mass plot for 2M1207b (red square), solar
    system planets and \bpic{} b (blue squares), and brown dwarfs
    (black circles, gray shade). Final rotation rates for 2M1207b and
    \bpic{} b estimated based on conservation of angular momentum are
    plotted with faint red and blue squares, respectively. The mass of brown dwarfs are assumed
    to be 30 \mjup{}. The gray rectangle that has a $\pm$15 \mjup
    range in $x$, and a $\pm \sigma$ of brown dwarf periods range in
    $y$, indicates a region where brown dwarfs most likely to appear
    in this diagram. Rotation period monotonically decreases with the
    increase of mass. {\em Right}: ratio of modulation amplitude in J
    and H band vs. spectral type for 2M1207b and brown dwarfs. The
    point for 2M1207b is shifted to +$x$ for half spectral type for
    clarification.  The colors of the points represent J$-$H
    magnitude, and the sizes of the points are proportional to the
    J-band modulation amplitudes. The gray dashed line is the result
    of a linear fit to these points.  Correlation of J- and
    H-band modulation amplitude ratio and spectral type is shown.}
 \label{fig:5}
\end{figure*}

@@

An important result of our study is the direct measurement of the
photometric modulation period of a directly imaged planetary-mass
object. We infer the rotation period of 2M1207b to be the same as
the period of photometric variation. Although horizontal winds can cause the
measured period to differ from true rotation period, it is
unlikely that the difference is greater than the uncertainty of the
measurement, since for Jupiter and Saturn, typical wind speeds are
more than one order of magnitude smaller than the equatorial rotation speeds.
% Our
% inferred rotation period of 2M1207b is typical among those of brown dwarfs
% \citep{Reiners2008}.
In the left panel of Figure~\ref{fig:5} we
compare the rotation period of 2M1207b to the solar system planets,
\bpic{} b the only other directly imaged planet with an estimated
period and measured \vsini, and field brown dwarfs from the study of
\cite[][]{Metchev2015}. \citet[][]{Snellen2014} measured \vsini{} for
\bpic{} b and demonstrated that it fits a trend defined by Solar
System planets in which more massive planets have faster rotation
rates. The interesting finding that \bpic{} b, an exoplanet that
formed in a protoplanetary disk, follows this trend suggests a
possibly connection between planet mass, initial angular momentum, and
formation in a disk.

>>>>

%%% new command %%%
\newcommand{\ima}{\texttt{ima} files }
\newcommand{\flt}{\texttt{flt} files }
\newcommand{\eps}{$\mathrm{e}^{-}/\mathrm{s}$}
\newcommand{\tinytim}{\textit{Tiny Tim}}
\newcommand{\bpic}{$\beta$ Pic}
\newcommand{\vsini}{$v\sin i$}
\newcommand{\mjup}{M$_{\mbox{Jup}}$}
\newcommand{\revise}[1]{\textit{\textbf{{\color{red}{#1}}}}}
\newcommand{\period}{$10.5^{+1.5}_{-0.8}$}
\begin{document}

@@

\begin{abstract}
  Rotational modulations of brown dwarfs have recently provided
  powerful constraints on the properties of ultra-cool atmospheres,
  including longitudinal and vertical cloud structures and cloud
  evolution. Furthermore, periodic light curves directly probe the
  rotational periods of ultra-cool objects.  We present here, for the
  first time, time-resolved high-precision photometric measurements of
  a planetary-mass companion, 2M1207b.  We observed the binary
  system with HST/WFC3 in two bands and with two spacecraft roll
  angles. Using point spread function-based photometry, we reach a
  nearly photon-noise limited accuracy for both the primary and the
  secondary. While the primary is consistent with a flat light curve,
  the secondary shows modulations that are clearly detected in the
  combined light curve as well as in different subsets of the data.
  The amplitudes are 1.45\% in the F125W and 0.92\% in the F160W
  filters, respectively. We find a consistent period of
\period hours and
  similar phases in both bands. The J- and H-band amplitude ratio of
  2M1207b is very similar to a field brown dwarf that has identical
  spectral type but different J-H color.  Importantly, our study also
  measures, for the first time, the rotation period for a directly 
  imaged planetary-mass companion to another star.
\end{abstract}

@@

\begin{figure*}
  \centering
  \plottwo{original}{subtracted}\\
  \includegraphics[width=0.32\textwidth]{TT_2M1207A}
  \includegraphics[width=0.32\textwidth]{TT_2M1207b}
  \includegraphics[width=0.32\textwidth]{TT_residual}
  \caption{Point spread function subtraction allows isolating the
    secondary and accurately measuring its brightness in our WFC3
    F160W images. {\em Upper row:} original (left) and primary
    subtracted (right) images. After subtraction of the primary PSF
    and correction map, 2M1207b is detected at a high significance
    level. \revise{{\em Lower row:} examples of \tinytim{} PSFs for
      the primary and the secondary, and the correction map.}
    % {\em Upper}: original image, the position of 2M1207
    % b is indicated by a white circle.  {\em Lower:} residual image --
    % after the subtraction of the hybrid PSF, 2M1207b is detected at a
    % high significance level.
  }
  \label{fig:1}
\end{figure*}

@@

Intensity modulations introduced by heterogeneous clouds can be
directly observed and studied via time resolved observation and
rotational mapping
\citep[e.g.][]{Apai2013,Buenzli2012,Buenzli2015,Burgasser2013,Radigan2012,Yang2015,Metchev2015,Heinze2015,
Biller2015}. These techniques isolate the effect of cloud
properties and obtained great success in determining the rotation
period and unveiling the structures of the atmosphere of brown dwarfs
. \citet{Kostov2013}
demonstrated that these techniques can be applied to directly imaged
exoplanets, too, allowing comparative studies of objects with
different surface gravities. However, high contrast magnifies the
challenges for directly imaged exoplanets and planetary-mass
companions to acquire high-precision light curves comparing to
\revise{isolated} brown
dwarfs.

2M1207b \citet{Chauvin2004} is the first directly imaged extra-solar
planetary-mass companion. \citet{Chauvin2005} and \citet{Song2006}
confirmed that 2M1207b and its host 2M1207A form a bound, co-moving
system. 2M1207A and b have an angular separation of $0.78''$, which
corresponds to a projected separation of 41.2 AU at a distance of 52.4
pc \citep[e.g.][]{Ducourant2008}. Combining 2M1207b's age and near
infrared luminosity with brown dwarf cooling models
\citep[e.g.][]{Baraffe2003}, the object's mass is estimated to be
2.3-4.8 M$_{\mathrm{Jup}}$ \citep{Barman2011b}. Even though a
circumsubstellar disk was discovered around 2M1207A
\citep[][]{Sterzik2004}, the high companion-to-host mass ratio and
large separation argue for binary-like gravitational fragmentation
formation \revise{\citep{Lodato2005,Mohanty2007}}.

@@

The discovery of additional planetary-mass companions with similarly
red colors \revise{(e.g. AB pic B, \cite{Chauvin2005b}, HR8799bcde,
  \cite{Marois2008a,Marois2010})}and apparent under-luminosity have
highlighted 2M1207b as a template of low gravity ultra-cool
atmospheres but as of now understanding the composition and structure
of clouds and their gravity-dependence remained elusive.

In this {\em Paper} we present the first, high-contrast, high-cadence,
high-precision, time-resolved {\em Hubble Space Telescope} (HST)
photometric time series of 2M1207b, a directly imaged planetary-mass
object. We successfully detect rotational modulation and measure the
amplitudes in two bands and determine the rotational period. These
observations probe the spatial heterogeneity and vertical structure of
clouds in planetary mass objects for the first time.

@@

We obtained direct images of the 2M1207A+b system on UT 2014 April 11
from 08:07:47 \revise{(JD 2456758.838738)} to 16:53:18 \revise{JD
  2456759.203681} using HST and its Wide Field Camera 3 \citep[WFC3,
pixel scale=$0.130$mas/pixel,
][]{Mackenty2008} in the frame of the HST Program GO-13418 (PI:
D. Apai). We acquired the observations in filters F125W
($\lambda_{\mathrm{pivot}}$
= 1245.9 nm, full width at half maximum (FWHM) = 301.5 nm) and F160W
($\lambda_{\mathrm{pivot}}$
1540.52, FWHM = 287.9 nm), roughly corresponding to the J and H
bands. We used the $256\times256$
pixels sub-array mode to avoid memory dumps during the observations.
In order to provide a near-continuous coverage for detecting
modulations we observed the 2M1207 system in 6 consecutive HST orbits,
obtaining data with \revise{maximum} cadence of 1.78 minutes over a
baseline of 8 hours and 40 minutes. The observations were interrupted
by 58-minute long Earth occultations every 94 minutes.

The observations applied space craft rolls each two orbits to allow
roll-subtraction of the primary \citep[e.g.][]{Song2006}. The
telescope roll angles for orbit 1, 3, and 5, and those for 2, 4, and 6
differed by $25^{\circ}$. At the separation of 2M1207b, this angle
difference corresponds to a displacement of $0.34''$, or 2.75 and 2.30
resolution elements in F125W and F160W, respectively.
%In each orbit we took 8 SPARS10 exposure sequence with NSAMP=10,
%alternating between F160W and F125W filters, with 2--3 identical
%exposures in each exposure sequence.
\revise{In each orbit, we took 8 SPARS 10 exposure sequences
  alternating between F125W and F160W filters. Each sequence contained
  2--3 identical exposures that had 10 non-destructive read-outs and total
  exposure time of 88.4 s} To improve sampling and reduce the risk
that the core of point spread function (PSF) is affected by bad
pixels, we applied a 4-point dither pattern with differential ``X/Y''
offsets of 1.375" in the detector frame, providing optimal
non-integral (half pixel) step of 10.5 and 8.5 pixels in F125W and
F160W, respectively. In total, we obtained 70 and 64 images with 10 in
F125W and F160W, respectively.

@@

 \begin{figure*}
  \centering
  \plotone{systematics}
  \caption{F125W (left) and F160W (right) light curves under different
    variability verification tests. Individual measurements are
    plotted with gray crosses. Photometric measurements of the same exposure
    sequence are binned, and binned photometry are plotted with points
    or squares. Best fitted sinusoids are plotted with solid
    lines. {\em Upper}: binned measurements taken in dithering
    position 1 and 3 (red points) and that taken in 2 and 4 (blue
    squares) are plotted with different symbols. They demonstrate same
    trend of modulation. In upper left panel, the green line a sinusoid
    fitted with all parameters freely varying, and the purple line is
    a sinusoid fitted with the period set the same as that of
    F125W. {\em Middle}: sinusoids fitted without using the
    data taken in Orbit \#1. These curves are almost identical to the
    curves plotted in upper panel. {\em Lower}: photometry measured
    with AFEM-added images and best fitted sinul curves. These
    points and curves are also almost identical to those plotted
    in the upper panel.}
  \label{fig:2}

\end{figure*}

@@

\label{Results}

@@

% We used a MC method to analyze the light curve, and provided amplitudes
% and periods as well as their uncertainties for both filters. We
% generated series of Gaussian distributed random noise with standard deviation
% same as the photon noise, added the noise to the original light curves, and
% fitted sinusoids to the newly combined light curves. We repeated these
% procedures  for 100,000 times and obtain the distributions of period
% and amplitude (Figure \ref{fig:4}).

\begin{figure*}
  \centering
\plotone{posterior_dist}
  \caption{\revise{Posterior distribution for amplitude and period for F125W
    (left) and F160W (right) light curves, with the data from the
    first orbit included (upper) and excluded (lower). In each panel,
    the univariate distributions for period and amplitude are plotted
    along side with the joint distributions. The measurements and
    uncertainties of period and amplitude are shown on the upper left
    corner of each panel.}}
  \label{fig:4}
\end{figure*}

\revise{To further constrain the amplitude and period of the light
  curve, we performed sinusoidal fit using Markov Chain Monte Carlo (MCMC)
  method.  The priors for the amplitude,
  the period and the phase of the sinusoid were all assumed to be
  uniform distributions. The ranges of the prior
  distributions were 0 to 5\% for the amplitude, 2 to 40 hours for the
  period, and 0 to $2\pi$ for the phase. The posterior distributions of
  the period and amplitude were very well constraint
  for both two filters so that we were able to measure the amplitudes
  and periods as well as their uncertainties. We also did the
fitting excluding data from the first orbit. In this case, posterior
distributions are not as well
constraint as those for including all the data, especially the
posterior distributions for period have significant proportion at the
long period end. However, for both two filters, the positions of the
peaks  of distributions of period and amplitude agree within 1-$\sigma$ uncertainty with or
without including data from the first orbit.}

\revise{We note that if the data from the first orbit were excluded, the
  light curves could be reasonably fitted to a linear trend. When ignoring
  data from the first orbit, For F125W light
  curve, the reduced  $\chi^{2}$ for sinusoidal and linear fits are
  1.138 and 1.143, and for F160W are 0.944 and 0.953. However, We are
  in favor of the sinusoidal model based on physical viability (See
  detailed discussion in \S\ref{sec:discussion}). }

\section{Result}

We present the first high-contrast, high-cadence, and high-precision
photometry of a directly imaged planet or planetary-mass
companion around another star. Our observations reveal a modulation in the light curve of
the $\sim 4 \mathrm{M_{{Jup}}}$ companion 2M1207b, the first detection
of modulations in directly imaged planetary-mass objects. The best
fit periods for F125W and F160W are $11.1_{-1.2}^{+2.0}$ and $9.3_{-0.8}^{+2.0}$ hours,
respectively. We also fitted the two band light curves together forcing the
periods of two sinusoids to be the same, and obtained a period
of $10.5^{+1.5}_{-0.8}$ hours.

@@

\section{Discussion}
\label{sec:discussion}
\begin{figure*}
  \centering
  \plottwo{rotationDiagram}{JH}
  \caption{comparison of 2M1207's rotation period and color change
    with brown dwarfs, \bpic{} b, and solar system planets. {\em
      Left}: period vs. mass plot for 2M1207b (red square), solar
    system planets and \bpic{} b (blue squares), field brown dwarfs
    from the study of \citet[][]{Metchev2015}
    (black circles, gray shade) and young brown dwarfs from the study
    of \citet{Scholz2015} (blue
    triangles). Final rotation rates for 2M1207b and \bpic{} b
    estimated based on conservation of angular momentum are plotted
    with faint red and blue squares, respectively. The mass of brown
    dwarfs are assumed to be $\sim30$ \mjup{}, and the difference mass
    of field brown dwarfs and young brown dwarfs are added
    artificially for the sake of clarity . The gray rectangle that has a
    $\pm$15
    \mjup range in $x$,
    and a $\pm \sigma$
    of field brown dwarf periods range in $y$,
    indicates a region where field brown dwarfs most likely to appear in
    this diagram. Rotation period monotonically decreases with the
    increase of mass. {\em Right}: ratio of modulation amplitude in J
    and H band vs. spectral type for 2M1207b and brown dwarfs. The
    point for 2M1207b is shifted to +$x$
    for half spectral type for clarification.  The colors of the
    points represent J$-$H
    magnitude, and the sizes of the points are proportional to the
    J-band modulation amplitudes. The gray dashed line is the result
    of a linear fit to these points.  Correlation of J- and H-band
    modulation amplitude ratio and spectral type is shown.}
 \label{fig:5}
\end{figure*}

@@

An important result of our study is the direct measurement of the
photometric modulation period of a directly imaged planetary-mass
object. \revise{Although a linear trend can also provide reasonable fit to
  the light curves when data from the first orbit are excluded, it is
  not a physical viable model. The light curve of a planetary mass
  companion is either a complete flat line, or a periodical curve with constant mean. On the contrary,
  2M1207b is not likely to constantly getting bright or faint. The
  flatness of the light curves of 2M1207A demonstrate that there is no
  instrumental sensitivity drift or any similar effect that can
  introduce a $\sim9$ hour slope. We do include very 
  long period sinusoid as possible solution for our MC fit, and
  demonstrate that a sinusoid with period of $\sim10$
  hour is the most probable solution with or without including data
  from the first orbit.} We infer the rotation period of
2M1207b to be the same as the period of photometric
variation. Although horizontal winds can cause the measured period to
differ from true rotation period, it is unlikely that the difference
is greater than the uncertainty of the measurement, since for Jupiter
and Saturn, typical wind speeds are more than one order of magnitude
smaller than the equatorial rotation speeds.
% Our
% inferred rotation period of 2M1207b is typical among those of brown dwarfs
% \citep{Reiners2008}.
In the left panel of Figure~\ref{fig:5} we compare the rotation period
of 2M1207b to the solar system planets, \bpic{} b the only other
directly imaged planet with an estimated period and measured \vsini,
field brown dwarfs from the study of \cite[][]{Metchev2015},
\revise{and young brown dwarfs (UpSco, age $\sim 11$ Myr) from
  \cite[][]{Scholz2015}}. \citet[][]{Snellen2014} measured \vsini{}
for \bpic{} b and demonstrated that it fits a trend defined by Solar
System planets in which more massive planets have faster rotation
rates. The interesting finding that \bpic{} b, an exoplanet that
formed in a protoplanetary disk, follows this trend suggests a
possibly connection between planet mass, initial angular momentum, and
formation in a disk.

Excitingly, our measurement of the rotation period of 2M1207b, a
planet mass companion that has similar age to \bpic{} b, has a
rotation period that fits in the same trend, as well as majority of
the brown dwarfs. As 2M1207b and \bpic{} b evolve and cool down, they are
expected to shrink to the size of Jupiter. Order of magnitude
estimation based on the conservation of angular momentum results in
final rotation periods of $\sim 5$ h and $\sim 3$ hours
\citep{Snellen2014} for 2M1207b and \bpic{} b, respectively, which
still fit to the period vs. mass trend. Although 2M1207A is known to
host a circumsubstellar disk \citep{Sterzik2004}, the low mass of
typical brown dwarf disks \citep[e.g.]{Klein2003, Mohanty2013} and its
large separation argue against the possibility that 2M1207b has formed
in a protoplanetary disk.  The result that objects formed in different
scenarios share the same trend of period vs. mass suggests that
rotation periods -- in absence of well-determined ages -- may
contribute insufficient evidence for a formation in a disk vs. in a
cloud core environment.

\revise{Rotation period of $\sim 10$
  hour is significantly larger than field brown dwarfs from the sample
  of \cite{Metchev2015}, but comparable to the median period
  of the sample of \cite{Scholz2015}, of which the age is similar
  to 2M1207b. The rotation period of 2M1207b
  is agreed within the range predicted by evolutionary track
  established from measurements of brown dwarfs assuming conservation
  of angular momentum (Fig. 4 in \cite{Scholz2015}), and much longer
  than the break-up limit, i.e. the rotation period where the
  equatorial centrifugal force exceeded gravitational force. Observed
  rotation rate for brown dwarfs of different ages show little
  evidence of angular momentum loss at the age of several Myr as
  oppose to low mass stars
  \citep{Bouvier2014, Scholz2015}, and agree with the model of solid
  body rotation for the fully convective interior. Rotation periods of
planetary mass objects with well established age measurement, can
place stronger constrain on the internal structure, since angular
momentum loss through disks is even less likely comparing with brown dwarfs.}
